---
title: "Tutorial work flow with objective functions"
author: "ACS"
output: 
  html_document:
      toc: true
      toc_float: true
      toc_depth: 4
      number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gcamland)
library(dplyr)
library(tidyr)
```

# Motivation/notes

- Operate on a truncated version of the `run_ensemble` function for testing and developing analysis work flow with objective functions:
`run_ensemble_no_analysis`. This version removes the `lpdf` and `lprior` arguments as well as the `if` statement that launches `run_bayes`. 
This seems like a cleaner way to support two different styles of analysis; otherwise, we'd have to add more arguments to `run_ensemble` to 
specify which objective functions to use, and unless a run was doing both bayesian and objective function analysis, only a subset of the 
arguments would ever be used. Seems cleaner to the do the run and break out the analysis after the fact, whether bayesian or objective function.
- This approach will rely on some of the utility functions defined in `bayesian.R` like `get_historical_land_data`. These have been
moved to their own script, and some have been updated from `internal` to `export` so that they can still be called from `bayesian.R` functions.



# Setup params

Pick the number of runs to do

```{r}
Nruns <- 2
```

# Do some runs 
Now, by default, no analysis is run. Only the ensemble of runs is generated, and analysis can be done after the fact. 

An example of the call to run the model, as well as the progress print statements included.
```{r}
set.seed(1)

scenObjects <- run_ensemble(N=Nruns)
```

# Run the objective function analysis

- Right now, the `run_objective` function evaluates user-specified objective function for each land type separately. Certainly _can_ give an argument with fewer land types.
- idea is that it's all so fast and not that big, just go ahead and calculate for each land type. Then when we call `MAP_objective` (which we should probably pick a different name for), _that_ function also takes a list of land types of interest, averages the objective function across those land types, and then picks the parameter set that gives you the smallest average objective function value (ie minimized error) across those land types.
- Also, currently, from `get_historical_land_data` in `utility-functions-analysis.R` (previously in `bayesian.R`), this only supports comparison to FAO GCAM commodities - mostly crop data and not Forest or anything. A note to change this is included in the documentation of `MAP_objective`.

- Note that land cover types X region combinations that have a standard deviation of 0 in the observational data, warnings will
be thrown and the metrics `nrms`, `centerednrms`, and `kge` will return NA values. 
- For land cover types X region combinations that have a standard deviation of 0 in the _modeled_ data, warnings will
be thrown and the metric `kge` will return NA values. In the United States, this means PalmFruit. 

An example of the function call to run the objective function analysis.
```{r}

scenObjectsEvaluated <- run_objective(scenObjects)
```

And different views of the results of that analysis. Note that the list item `$mObjFunEval` for each scenario is where this analysis is contained:
```{r}

str(scenObjectsEvaluated)

names(scenObjectsEvaluated[[1]])

knitr::kable(head(scenObjectsEvaluated[[1]]$mObjFunEval))

knitr::kable(tail(scenObjectsEvaluated[[1]]$mObjFunEval))

knitr::kable(head(scenObjectsEvaluated[[1]]$mObjFunEval %>%
                    filter(land.type == 'PalmFruit', 
                           objfun == 'kge')))

```

# Organize these evaluated results 

Using the `grand_table_objective` function results of the objective function evaluations 
are reorganized into a grand table of parameters.
The table produced includes the model parameters and all user-specified objective
function evaluations, for all of the models in the input list.

An example of the call to this function and a quick look at the structure of the results.
```{r}
grand_table_objective(aScenarioList = scenObjectsEvaluated) %>%
  # these lines of code are because the `left_join` in 
  # `add_parameter_data` (from utility-functions-analysis.R) is only
  # a join on scenario, rather than scenario and region.
  # Unclear if should update that join or if we somehow want to keep it
  # separate; ideally will figure out more when do other regions.
  # For now, leaving it alone so that the `grand_table` works as it
  # did before. 
  # (relevant because both `grand_table` and `grand_table_objective`
  # call `add_parameter_data`)
  filter(region.x == region.y) %>%
  rename(region = region.x) %>%
  select(-region.y) ->
  GTobjective

knitr::kable(head(GTobjective))
```


And a view of a reshaped, 'prettier' (that is, wide) version of these results. 
However, the long version `GTobjective` is what the next step in analysis is expecting.
```{r}

GTobjective %>%
  spread(objfun, objfunval) ->
  GTobjective_pretty

knitr::kable(head(GTobjective_pretty))
```


# Select the parameter set from runs that minimize specified obj fcn across specified land types

`Map_objective` is a minimizer function. Itcans the runs in a given sample of
model runs and selects the parameter set from those runs that minimizes the
mean across user-specified `landtypes` of a user-specified objective function, 
`objfun_to_min`. 

## Exceptions

For `bias`, the quantity being minimized is actually the average over specified
land types of `abs(bias)`, as a bias of 0 is perfect performance, and a bias of
0.25 vs -0.25 are equally 'good'.

For `kge`, the quantity being minimized is actually the average over specified
land types of `1 -kge`. This is chosen because of the following:

1. Values of `kge` can range anywhere from -infinity to +1, with a value of +1
being perfect model performance, therefore `kge` is not directly useful as a 
quantity to be minimized.

2. Values of `1-kge` range from 0 to +infinity, with 0 corresponding to perfect
model performance, as well as being the minimum value achieved by this measure. 
Additionally, because all values of this quantity are positive, no cancelation 
of errors may occur when calculating the mean across specified land types.

Further details are given in the documentation of the `MAP_objective` function.

## default arguments

Default arguments of `MAP_objective` currently is focused on RMS 
(`objfun_to_min = 'rms'`) for all GCAM commodities 
`landtypes = c( "Corn", "FiberCrop", "MiscCrop", "OilCrop", "OtherGrain", "PalmFruit",  "Rice", "Root_Tuber", "SugarCrop",  "Wheat")` 
that we run and have FAO comparison data for. 

`MAP_objective` is taking the average of RMS across all of these land types, 
and selecting the parameter set that minimizes that quantity.

Currently no support for mixing and matching different objective functions 
for different crops. Seems doable but not implemented yet (eg maybe you want
parameters that minimize both Corn's bias and OtherGrain's KGE.)


An example call to the `MAP_objective` function assuming default arguments, 
and a printout of the results:
```{r}

knitr::kable(MAP_objective(GTobjective))
```


## Corn, OilCrop bias example

Select parameters that minimize the average (absolute value of) bias in Corn
and OilCrop.

The current implementation does not allow for cancellation of errors: the 
average across specified land types of `abs(bias)` is what is actually being
minimized in `MAP_objective`. So, for example, a positive bias in Corn and a
negative bias in OilCrop will _not_ cancel out to give an artifically small
mean error.


```{r}

knitr::kable(MAP_objective(GTobjective, objfun_to_min = 'bias', landtypes = c('Corn', 'OilCrop')))
```

### NOTE
The error reported in the column `landTypeMeanObjFunVal` is what you think: it's the average error only across the specified land types. If you want to see what the error on Wheat is for this example, the results from `grand_table_objective` (ie `GTobjective`) must be filtered and examined.
```{r}
GTobjective %>%
   filter(land.type == 'Wheat')
```

## OtherGrain rms example

```{r}

knitr::kable(MAP_objective(GTobjective, objfun_to_min = 'rms', landtypes = c('OtherGrain')))
```


